Sampling requirements to detect ecosystem change

### 2020-02-05 R/manuscript_draft.Rmd
added pieces to theoretical results section

We began with a theoretical exploration of the sampling requirements to detect a changepoint. We modeled a simple first order autoregressive (AR-1) process with a response variable that represents either population size, biodiversity, or some metric of community composition. The model includes temporal autocorrelation ($\phi$), the mean of the process ($\mu_X$), and a white noise term ($w_t$). The white noise term is a normal distribution with mean ($\mu_w$) and variance ($\sigma$).

@@

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width = 7.2, fig.height= 8}
library(ggplot2)
library(gridExtra)
p1 <- ggplot(myoutput[myoutput$target_cpt==1,], aes(final_n,diff_real, pch=factor(n2_init))) + 
  geom_hline(aes(yintercept=0), col=adjustcolor("black", alpha.f = .4)) + 
  geom_point(alpha = .3) +
  facet_wrap(~method_f) + theme_bw() +
  labs(title="a. First changepoint",x="", y="Distance to real changepoint") +
  theme(legend.position = c(0.93, 0.6),
        legend.background = element_rect(fill = "#ffffffaa", colour = NA),
        legend.title=element_text(size=7), 
        legend.text=element_text(size=8)) +
  guides(pch=guide_legend(title="Initial number\nof sample"))
p2 <- ggplot(myoutput[myoutput$target_cpt==2,], aes(final_n,diff_real, pch=factor(n2_init))) + 
  geom_hline(aes(yintercept=0), col=adjustcolor("black", alpha.f = .4)) + 
  geom_point(alpha = .3) +
  facet_wrap(~method_f) + theme_bw() +
  labs(title="b. Second changepoint",x="", y="Distance to real changepoint") +
  theme(legend.position = c(0.93, 0.6),
        legend.background = element_rect(fill = "#ffffffaa", colour = NA),
        legend.title=element_text(size=7), 
        legend.text=element_text(size=8)) +
  guides(pch=guide_legend(title="Initial number\nof sample"))
p3 <- ggplot(myoutput[myoutput$target_cpt==3,], aes(final_n,diff_real, pch=factor(n2_init))) + 
  geom_hline(aes(yintercept=0), col=adjustcolor("black", alpha.f = .4)) + 
  geom_point(alpha = .3) +
  facet_wrap(~method_f) + theme_bw() +
  labs(title="c. Third changepoint",x="", y="Distance to real changepoint") +
  theme(legend.position = c(0.93, 0.6),
        legend.background = element_rect(fill = "#ffffffaa", colour = NA),
        legend.title=element_text(size=7), 
        legend.text=element_text(size=8)) +
  guides(pch=guide_legend(title="Initial number\nof sample"))
grid.arrange(p1,p2,p3,nrow=3)
```

>>>>

We began with a theoretical exploration of the sampling requirements to detect a changepoint. We modeled a simple first order autoregressive (AR-1) process (the discrete-time version of the Ornstein–Uhlenbeck process) with a response variable that represents either population size, biodiversity, or some metric of community composition. The model includes temporal autocorrelation ($\phi$), the mean of the process ($\mu_X$), and a white noise term ($w_t$). The white noise term is a normal distribution with mean ($\mu_w$) and variance ($\sigma$).

@@

```{r theoretical_results_min_time, fig.cap='Theoretical results here detailing parameter sensitivity - we should make this about sample size instead of probability correct',message=F,warning=F,echo=F,eval=T}

source('R/sampling_regular.R')
source('R/build_time_series.R')
source('R/updated_param_sensitivity.R')
source('R/calc_min_time2.R')

par(mfrow=c(1,3))
# Sigma plot
sigma_vec = seq(0.1,20,1)
min_time=NULL

for (param_num in 1:length(sigma_vec)){
  min_time = c(min_time,calc_min_time('sigma',sigma_vec[param_num]))
  print(min_time)
}

plot(sigma_vec,min_time,ylab='Number of samples required',xlab='sigma',pch=16,cex.lab=1.3,las=1,ylim=c(0,25),col=1)

phi_vec = seq(-1,1,0.2)
min_time=NULL

for (param_num in 1:length(phi_vec)){
  min_time = c(min_time,calc_min_time('phi',phi_vec[param_num]))
}

plot(phi_vec,min_time,ylab='Number of samples required',xlab='phi',pch=16,cex.lab=1.3,las=1,ylim=c(0,1),col=1)

# Shift size

shift_size_vec = seq(0.05,0.95,0.1)
min_time=NULL

for (param_num in 1:length(shift_size_vec)){
  min_time = c(min_time,calc_min_time('shift_size',shift_size_vec[param_num]))
}

plot(shift_size_vec,min_time,ylab='Number of samples required',xlab='shift size (percent)',pch=16,cex.lab=1.3,las=1,ylim=c(0,1),col=1)

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(gridExtra)
p1 <- ggplot(myoutput, aes(final_n,diff_real)) + 
  geom_point(alpha = .3) +
  facet_wrap(~method_f) + theme_bw() +
  labs(title="a",x="", y="Distance to real changepoint")
p2 <- ggplot(myoutput, aes(rep(0, nrow(myoutput)),diff_real))  +
  facet_wrap(~method_f) + 
  geom_violin() +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=.3) + 
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + theme_bw() +
  labs(title="b",x="", y="Distance to real changepoint")
grid.arrange(p1, p2, nrow = 2)
```
